---
title: 'Statistical Inference Project Part 2: Tooth Growth Analysis'
author: "Ryan P"
date: "4/8/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(data.table)
library(ggplot2)
library(datasets)
tg <- ToothGrowth
View(tg)
summary(tg)
str(tg)
```
```{r}
#Visualisation of Tooth Length by Dosage Frequency
g <- ggplot(aes(x=as.factor(dose), y=len), data = tg)
g <- g + labs(x= "Dose Frequency", y = "Growth", title = "Tooth Growth by
              Dosage Frequency")
g <- g + geom_point(aes(color = supp))
g
```
```{r}
b <- ggplot(data = tg, aes(x = as.factor(dose), y = len, fill = supp))
b <- b + geom_bar(stat = "identity")
b <- b + facet_grid(.~supp)
b <- b + labs (x= "Dose Frequency", y = "Growth") + guides(fill = guide_legend(
  title = "Supplement"))
b
```
Just from looking at the data visually, it appears there is little difference 
between supplements on how much of an effect they have. ** Analyze this***

The greatest effect on tooth length appears to be Dose Frequency
and not whether or not the Dose is supplemented with either OJ or VC.

Is there a significant difference between 1 dose per day and OJ and VC 
administered supplements? *** How do I analyze this? ***

There looks to be significantly increased growth within supplements with increased
dose frequency*** Analyze this***
Null Hypothesis One: Supplements have no effect on tooth growth.
Alternate One: Supplements increase tooth growth
```{r}


Null Hypothesis TWO: There is no difference on tooth growth between supplements.
Alternate Two: VC or OJ have different effects on tooth growth
Alt 2a: At does 0.5 mg
Alt 2b: At dosage 1.0 mg
Alt 2c: At dosage 2.0 mg


Null Hypothesis THREE: Dose Frequency has no effect on tooth growth
Alternative THREE: Dose Frequency affects tooth growth