---
title: 'Statistical Inference Project Part 2: Tooth Growth Analysis'
author: "Ryan Paulk"
date: "4/8/2021"
output: pdf_document
---
# Synopsis

The tooth growth data set is the resulting data collected for guinea pigs on a
modified diet and the resultant tooth growth. The two independent variables are the
dosage of Vitamin C (0.5, 1, and 2 mg) and the delivery supplement (orange juice,
"OJ", and ascorbic acid, "VC".) It is a small data set and can be preliminarily
explored with the View() function.

# Key Assumptions of the Data
1. Data is independent and identically distributed
2. Variances are considered to be unequal
3. Data is representative of the population
4. The t-distribution can correct for and normalize the small data set

# Data Exploration
(See Appendix for code)
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(data.table)
library(ggplot2)
library(datasets)
tg <- ToothGrowth
View(tg)
summary(tg)
str(tg)
```
Two graphs were produced to explore possible relationships.

```{r Tooth Growth Graph, fig.width = 5, fig.height = 3, echo = FALSE}
#Visualisation of Tooth Length by Dosage
g <- ggplot(aes(x=as.factor(dose), y=len), data = tg)
g <- g + labs(x= "Dose", y = "Growth", title = "Tooth Growth by
              Dosage")
g <- g + geom_point(aes(color = supp))
g
```

In the first graph, the effects of the two supplements appear to become less 
distinct with increased dose.

```{r Dosage Graph, fig.width = 5, fig.height = 3, echo = FALSE}
#Visualisation of Tooth Growth by Supplement
b <- ggplot(data = tg, aes(x = as.factor(dose), y = len, fill = supp))
b <- b + geom_bar(stat = "identity")
b <- b + facet_grid(.~supp)
b <- b + labs (x= "Dose", y = "Growth") + guides(fill = guide_legend(
  title = "Supplement"))
b
```

From the second graph, there is an apparent trend of increased tooth growth
with increased vitamin C dose, no matter which supplement is used.
It is still not clear if one supplement is distinctly better.

# Statistical Analysis

Three hypotheses will be investigated using t.tests and the resulting p-values
will be discussed.

# Hypothesis One: Supplements have different effects on tooth growth
# Null Hypothesis: There is no significant difference between Supplements

```{r Supplements, echo = FALSE}
OJ <- tg$len[tg$supp == "OJ"]
VC <- tg$len[tg$supp == "VC"]
```
```{r}
t.test(OJ, VC, paired = FALSE)$p.value

```
The null hypothesis fails to be rejected and indicates that there is no 
difference between supplements on tooth growth.

# Hypothesis Two: Size of dose has an effect on tooth growth
# Null Hypothesis: Size of dose has no effect on tooth growth
```{r Doses, echo = FALSE}
dose.5 <- tg$len[tg$dose == 0.5]
dose1 <- tg$len[tg$dose == 1]
dose2 <- tg$len[tg$dose == 2]
#Default confidence interval is 0.95
```
```{r}
t.test(dose.5, dose1, alternative = "less", paired = FALSE)$p.value
```
```{r}
t.test(dose1, dose2, alternative = "less", paired = FALSE)$p.value
```
The p-values returned allow us to reject the null hypothesis and conclude that 
the increased tooth growth may be related to increases in dose.


# Hypothesis Three: One supplement's delivery method has a greater effect on tooth growth than the other
# Null Hypothesis Two: There is no difference between supplements' effects on tooth growth

```{r Supplement Doses, echo = FALSE}
# Supplement delivery method and doses defined
oj.5 <- tg$len[tg$dose == 0.5 & tg$supp == "OJ"]
oj1 <- tg$len[tg$dose == 1 & tg$supp == "OJ"]
oj2 <- tg$len[tg$dose == 2 & tg$supp == "OJ"]
vc.5 <- tg$len[tg$dose == 0.5 & tg$supp == "VC"]
vc1 <- tg$len[tg$dose == 1 & tg$supp == "VC"]
vc2 <- tg$len[tg$dose == 2 & tg$supp == "VC"]
```
First, the smallest doses were compared.
```{r}
t.test(oj.5, vc.5, alternative = "greater", paired = FALSE)$p.value
```
The p-value from this one-sided t-test indicates that OJ has a greater effect
on tooth growth than ascorbic acid at the lowest dose.

```{r}
t.test(oj1, vc1, alternative = "greater", paired = FALSE)$p.value
```
Again, the same t-test with medium doses shows OJ having a stronger effect.

```{r}
t.test(oj2, vc2, paired = FALSE)$p.value
```
Finally, at the highest dose and in agreement to what was seen in the graphs,
there is no significant difference between supplements at the highest dose.

# Conclusion

From the data we can conclude that it does not matter which delivery method or 
supplement is used so long as higher vitamin C doses are  made available for 
optimal tooth growth.

\pagebreak

# Appendix
```{r ref.label = "setup", eval = FALSE}
```
```{r ref.label = "Tooth Growth Graph", eval = FALSE}
```
```{r ref.label = "Dosage Graph", eval = FALSE}
```
```{r ref.label = "Supplements", eval = FALSE}
```
```{r ref.label = "Doses", eval = FALSE}
```
```{r ref.label = "Supplement Doses", eval = FALSE}
```


```{r}
OJ <- tg$len[tg$supp == "OJ"]
VC <- tg$len[tg$supp == "VC"]
t.test(OJ, VC, paired = FALSE)$p.value

```