---
title: 'Statistical Inference Project Part 2: Tooth Growth Analysis'
author: "Ryan P"
date: "4/8/2021"
output: pdf_document
---
The tooth growth data set is the resulting data set of an modified diet for
guinea pigs and the resultant tooth growth. The two independent variables are the
dosage of Vitamin C (0.5, 1, and 2 mg) and the delivery supplement (orange juice,
"OJ", and ascorbic acid, "VC".) It is a small data set and can be preliminarily
explored with the View() function.
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(data.table)
library(ggplot2)
library(datasets)
tg <- ToothGrowth
View(tg)
summary(tg)
str(tg)
```
Possible trends and effects can be better realized through graphs.
```{r}
#Visualisation of Tooth Length by Dosage Frequency
g <- ggplot(aes(x=as.factor(dose), y=len), data = tg)
g <- g + labs(x= "Dose Frequency", y = "Growth", title = "Tooth Growth by
              Dosage Frequency")
g <- g + geom_point(aes(color = supp))
g
```
Visually, the data shows there might be significant differences between OJ and 
ascorbic acid at lower doses, but none at higher.
```{r}
b <- ggplot(data = tg, aes(x = as.factor(dose), y = len, fill = supp))
b <- b + geom_bar(stat = "identity")
b <- b + facet_grid(.~supp)
b <- b + labs (x= "Dose Frequency", y = "Growth") + guides(fill = guide_legend(
  title = "Supplement"))
b
```
Here, just from looking at the bar graphs, it seems apparent that increased
dosages for both OJ and ascorbic acid increase tooth growth.

Just from looking at the data visually, it appears there is little difference 
between supplements on how much of an effect they have. ** Analyze this***

The greatest effect on tooth length appears to be Dose Frequency
and not whether or not the Dose is supplemented with either OJ or VC.

Is there a significant difference between 1 dose per day and OJ and VC 
administered supplements? *** How do I analyze this? ***

There looks to be significantly increased growth within supplements with increased
dose frequency*** Analyze this***
Null Hypothesis One: Supplements have no effect on tooth growth.
Alternate One: Supplements increase tooth growth
# There is no data for no supplement so this hypothesis can not be tested.
```{r}
### Variances are all assumed to be unequal, var.equal = FALSE
```
Null Hypothesis TWO: There is no difference on tooth growth between supplements.
Alternate Two: VC or OJ have different effects on tooth growth
Alt 2a: At does 0.5 mg
Alt 2b: At dosage 1.0 mg
Alt 2c: At dosage 2.0 mg
```{r}
oj.5 <- tg$len[tg$dose == 0.5 & tg$supp == "OJ"]
oj1 <- tg$len[tg$dose == 1 & tg$supp == "OJ"]
oj2 <- tg$len[tg$dose == 2 & tg$supp == "OJ"]
vc.5 <- tg$len[tg$dose == 0.5 & tg$supp == "VC"]
vc1 <- tg$len[tg$dose == 1 & tg$supp == "VC"]
vc2 <- tg$len[tg$dose == 2 & tg$supp == "VC"]
```
With these defined, we can test between supplements at different dosages.
```{r}
t.test(oj.5, vc.5, alternative = "greater", paired = FALSE)
```
The significance of this result indicates that OJ has a greater effect on 
tooth growth than ascorbic acid.

```{r}
t.test(oj1, vc1, alternative = "greater", paired = FALSE)
```
This again indicates that OJ has a greater effect on tooth growth than ascorbic
acid.
```{r}
t.test(oj2, vc2, paired = FALSE)
```
At higher doses there is no significant difference between the supplements'
effects on tooth growth.

#Null Hypothesis THREE: Dose Frequency has no effect on tooth growth
#Alternative THREE: Dose Frequency affects tooth growth
```{r}
dose.5 <- tg$len[tg$dose == 0.5]
dose1 <- tg$len[tg$dose == 1]
dose2 <- tg$len[tg$dose == 2]
#Default confidence interval is 0.95
```
```{r}
t.test(dose.5, dose1, alternative = "less", paired = FALSE)
```
```{r}
t.test(dose1, dose2, alternative = "less", paired = FALSE)
```
These tests show that there is a significant increase in tooth growth with 
increased dosages.