---
title: "Statistical Inference Project Part 1: Simulation Exercise"
author: "Ryan P"
date: "4/7/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
library(dplyr)
library(data.table)
library(ggplot2)
```
This project sought to compare the exponential distribution in R with the
Central Limit Theorem and demonstrate their similarity. The exponential
distribution was simulated forty times (n = 40) at a rate of lambda = 0.2.
The mean of the exponential distribution is defined as 1/lambda and the 
standard deviation also at 1/lambda. This simulation was repeated 1000
times and graphed below.
```{r}
# Set seed to make data replicable
set.seed(4321) 
#define parameters
lam <- 0.2
mu <- 1/lam
stdev <- 1/lam
n <- 40
nsim <- 1000
# Generate data matrix of simulated means and graph
simdata <- matrix(rexp(n*nsim, rate = lam), nsim)
simmeans <- rowMeans(simdata)
hist(simmeans, main = "Frequency Distribution of Simulated Means", xlab = "Simulated Mean")
```
It appears fairly normal centered about five.
The sample mean was determined by taking the mean of the simulated means.

```{r}
#Show the sample mean and compare it to the theoretical
#Sample mean:
smean <- mean(simmeans)
smean
```
Next the theoretical mean of the exponential distribution is defined as mu = 1/lambda.

```{r}
#Theoretical mean of exponential distribution is defined as mu = 1/lambda
#Theoretical mean:
them <- 1/lam
them
```
And they appear to be very close.

Next, variance for the simulation was calculated as the sample variance
of the simulated means.


```{r}
#Show how variable the sample is (via variance) and compare it to the theoretical variance of the distribution
#Sample variance
svar <- var(simmeans)
svar
```
And the theoretical variance was calculated with the sample size (n) factored in so
that it was equal to 1/(n*lambda^2).
```{r}
#Theoretical variance
#Theoretical variance of exponential dist = 1/(n*lambda^2) and 
tvar <- 1/(n*lam^2)
tvar
```
Sample standard deviations were then calculated.
```{r}

#Standard deviations
#Sample Std Dev
sstd <- sd(simmeans)
sstd
```
Again, taking into account the sample size when calculating the theoretical
standard deviation.
```{r}
#Theoretical Std Dev
tstd <- 1/(lam*sqrt(n))
tstd
```
All of the sample's attributes appear to be close to those of the theoretical 
distribution.


To show that the sample of the exponential distribution is normal, the data was
plotted and fitted with both the sample's curve and the theoretical curve. The
means were then marked with vertical lines that overlapped.
```{r}
#Show distribution is ~N
samples <- data.frame(simmeans)
g <- ggplot(samples, aes(x=simmeans))
g <- g + geom_histogram(aes(y = ..density..), fill = "white", color = 17)
g <- g + geom_vline(xintercept = smean, color = 3 )
g <- g + geom_vline(xintercept = them, color = "red")
g <- g + labs(x = "Sample Mean", y = "Density", title = "Density Distrubtion of Sample Means")
g <- g + stat_function(fun = dnorm, args = list(mean = smean, sd = sstd), color= 3)
g <- g + stat_function(fun = dnorm, args = list(mean = them, sd = tstd), color = "red")
g 
```

```{r}
#Confidence Intervals maybe?
#Sample CI
sCI <- smean + c(-1, 1)*1.96*sstd/sqrt(n)
sCI
#Theoretical CI
tCI <- them + c(-1, 1)+1.96*tstd/sqrt(n)
tCI
```