---
title: "Statistical Inference Project Part 1: Simulation Exercise"
author: "Ryan Paulk"
date: "4/7/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
library(dplyr)
library(data.table)
library(ggplot2)
```
# Synopsis
This project sought to compare the exponential distribution in R with the
Central Limit Theorem and demonstrate their similarity. The exponential
distribution was simulated forty times (n = 40) at a rate of lambda = 0.2.
The mean of the exponential distribution is defined as 1/lambda and the 
standard deviation also at 1/lambda. This simulation was repeated 1000
times and graphed below.
```{r fig.height = 2.5, fig.width = 4.5}
# Set seed to make data replicable
set.seed(4321) 
#define parameters
lam <- 0.2
mu <- 1/lam
stdev <- 1/lam
n <- 40
nsim <- 1000
# Generate data matrix of simulated means and graph
simdata <- matrix(rexp(n*nsim, rate = lam), nsim)
simmeans <- rowMeans(simdata)
hist(simmeans, main = "Frequency Distribution of Simulated Means", xlab =
       "Simulated Mean")
```

The distribution appears to be Gaussian which is a tell-tale sign that 
it is distributed Normally.

The sample mean was identified by taking the mean of the simulated means.
```{r}
#calculate and store the sample mean of the simulated means
smean <- mean(simmeans)
smean
```
For comparison, the theoretical mean of the exponential distribution is defined
as mu = 1/lambda.

```{r}
#calculate and store the theoretical mean
them <- 1/lam
them
```
And they appear to be very close.

Next, variance for the simulation was calculated as the sample variance
of the simulated means.
```{r}
#Sample variance
svar <- var(simmeans)
svar
```
And the theoretical variance was calculated with the sample size (n) factored
in sothat it was equal to 1/(n*lambda^2).
```{r}
#Theoretical variance
tvar <- 1/(n*lam^2)
tvar
```
Sample standard deviations were then calculated.
```{r}
#Sample standard deviation
sstd <- sd(simmeans)
sstd
```
Again, taking into account the sample size when calculating the theoretical 
standard deviation.
```{r}
#Theoretical standard deviation
tstd <- 1/(lam*sqrt(n))
tstd
```
Furthermore, a comparison of 95% confidence intervals shows little deviation 
from the theoretical distribution.
```{r}
#Sample CI
sCI <- smean + c(-1, 1)*1.96*sstd/sqrt(n)
sCI
#Theoretical CI
tCI <- them + c(-1, 1)*1.96*tstd/sqrt(n)
tCI
```
All of the sample's attributes appear to be close to those of the theoretical 
distribution.

To show that the sample of the exponential distribution is normal, the data was
plotted and fitted with both the sample's curve and the theoretical curve. The
means were then marked with vertical lines that overlapped.
```{r fig.height = 2.5, fig.width = 4.5}
#Show distribution is ~N
samples <- data.frame(simmeans)
g <- ggplot(samples, aes(x=simmeans))
g <- g + geom_histogram(aes(y = ..density..), fill = "white", color = 17)
g <- g + geom_vline(xintercept = smean, color = 3 )
g <- g + geom_vline(xintercept = them, color = "red")
g <- g + labs(x = "Sample Mean", y = "Density", title = "Density Distrubtion of
              Sample Means")
g <- g + stat_function(fun = dnorm, args = list(mean = smean, sd = sstd), 
                       color= 3)
g <- g + stat_function(fun = dnorm, args = list(mean = them, sd = tstd), 
                       color = "red")
g 
```

# Conclusion
Based on the simulation's distribution's similarity to the theoretical 
distribution it can be said to behave as predicted by the Central Limit Theorem.